## depobio stats

GBIF France publishes a large group of data from a nationally mandated repository for EIA data called **depobio**. 

The relevant datasets are all tagged with the keyword “depobio”. 

https://www.gbif.org/dataset/search?offset=4000&q=depobio&type=OCCURRENCE

```{r setup, include=FALSE}
library(dplyr)
library(purrr)
hh <- 10
```

## total occurrence counts

```{r total occ records, echo=FALSE,include=FALSE}
total_occ_counts = "https://api.gbif.org/v1/dataset/search/export?format=TSV&q=depobio&type=OCCURRENCE" %>%
readr::read_tsv() %>%
head(hh) %>% 
pull(occurrence_records_count) %>% 
sum()
```

The total number of occurrence records published under the keyword "depobio":

`r as.character(total_occ_counts)`

## percent total by year (eventDate)

```{r percent total by event date, echo=FALSE,include=FALSE}

occ_by_year = "https://api.gbif.org/v1/dataset/search/export?format=TSV&q=depobio&type=OCCURRENCE" %>%
readr::read_tsv() %>% 
head(hh) %>% 
select(dataset_key) %>%
mutate(facet = map(dataset_key,~rgbif::occ_search(datasetKey=.x,facet = c("year"),limit=0,year.facetLimit = 100000)$facets$year)) %>% 
tidyr::unnest(cols="facet") %>%
group_by(name) %>% 
summarise(n_occ_depobio = sum(count)) %>% 
glimpse() %>% 
merge(rgbif::occ_search(country="FR",facet = c("year"),limit=0,year.facetLimit = 100000)$facets$year,by="name") %>%
select(year=name,n_occ_depobio,n_occ_fr=count) %>% 
mutate(per_occ_depobio = (n_occ_depobio/n_occ_fr)*100) %>% 
mutate(year = as.numeric(year)) %>%
arrange(year)  

```

`r occ_by_year %>% knitr::kable()`

## number of unique species

```{r n_unique_species, echo=FALSE,include=FALSE}

unique_depobio_species = "https://api.gbif.org/v1/dataset/search/export?format=TSV&q=depobio&type=OCCURRENCE" %>%
readr::read_tsv() %>%
head(hh) %>%
glimpse() %>%
pull(dataset_key) %>% 
map(~
rgbif::occ_search(datasetKey=.x,facet = c("speciesKey"),limit=0,speciesKey.facetLimit = 100000)$facets$speciesKey
) %>% 
compact() %>%
bind_rows() %>% 
unique() %>%
group_by(name) %>%
summarise(count = sum(count)) %>% 
glimpse() 

```

Total number of unique species for datasets published under the keyword "depobio":

`r nrow(unique_depobio_species)`

<!-- ## number unique species, but not in any other dataset in France -->

<!-- ```{r not_in_previously_france, echo=FALSE,include=FALSE} -->

<!-- # depobio_dataset_key = "https://api.gbif.org/v1/dataset/search/export?format=TSV&q=depobio&type=OCCURRENCE" %>% -->
<!-- # readr::read_tsv() %>% -->
<!-- # head(hh) %>% -->
<!-- # glimpse() %>% -->
<!-- # pull(dataset_key)  -->
<!-- #  -->
<!-- # unique_france_species = rgbif::occ_search(country="FR",facet = c("datasetKey"),limit=0,datasetKey.facetLimit = 100000)$facets$datasetKey %>%  -->
<!-- # filter(!name %in% depobio_dataset_key) %>%  -->
<!-- # pull(name) %>% -->
<!-- # map(~ -->
<!-- # rgbif::occ_search(datasetKey=.x,facet = c("speciesKey"),limit=0,speciesKey.facetLimit = 100000)$facets$speciesKey  -->
<!-- # ) %>%  -->
<!-- # compact() %>% -->
<!-- # map(~ .x$name) %>%  -->
<!-- # flatten_chr() %>%  -->
<!-- # unique()  -->

<!-- ``` -->

<!-- These species only occur in datasets published using: -->

<!-- `r length(unique_depobio_species$name[!unique_depobio_species$name %in% unique_france_species])` -->

<!-- # ```{r unique_in_france_table, echo=FALSE,include=TRUE,results='asis'} -->
<!-- #  -->
<!-- # unique_depobio_species$name[!unique_depobio_species$name %in% unique_france_species] %>% -->
<!-- # data.frame() %>% -->
<!-- # knitr::kable() -->
<!-- #  -->
<!-- # ``` -->

## iucn counts

```{r iucn_counts, echo=FALSE,include=FALSE}

unique_depobio_species_iucn = unique_depobio_species %>%
mutate(iucn_cat = map_chr(name,~try(rgbif::name_usage(key=.x,data="iucnRedListCategory")$data$category))) %>%
glimpse() 
```

```{r iucn_counts_print_table, echo=FALSE,include=TRUE,results='asis'}

unique_depobio_species_iucn %>%
group_by(iucn_cat) %>%
summarise(n_occ = sum(count),
n_species = n_distinct(name)
) %>%
arrange(-n_occ) %>% 
knitr::kable()

```

<!-- ## iucn counts unique to depobio in FR -->

<!-- ```{r iucn_counts_print_table_unique_FR, echo=FALSE,include=TRUE,results='asis'} -->

<!-- unique_depobio_species_iucn %>% -->
<!-- filter(!name %in% unique_france_species) %>% -->
<!-- group_by(iucn_cat) %>% -->
<!-- summarise(n_occ = sum(count), -->
<!-- n_species = n_distinct(name) -->
<!-- ) %>% -->
<!-- arrange(-n_occ) %>%  -->
<!-- knitr::kable() -->

<!-- ``` -->

## unique species by higher level taxa from schema used in country report

```{r htax, echo=FALSE,include=FALSE}

unique_depobio_species_htax = 
unique_depobio_species %>%
mutate(name_data=map(name,~rgbif::name_usage(key=.x)$data)) %>%
tidyr::unnest(cols="name_data")
```

```{r country_report, echo=FALSE,include=TRUE,results='asis'}
bony_fish =
c("1103","1104","494","495","1105","496","497","1106","498","499","1107","537","538","1153","547","1162","548","549","550","1163","1164","1165","1166","1167","1305","1067","1306","1307","1308","1068","1069","587","1310","588","589","1311","1312","1313","590","890","774","889","773","772","888","765","879","7191147","1108","1448","8454030","716","8510645","8706725","8602104","7192402","839","721","1446","1447","1492","8481794","7191407","723","7192754","1493","8454707","8617753","1444","10833565","8705315","8708973","7190953","7190987","7191426","724","7191588","7192755","1445","7192749","8603836","7190978","10726067","1450","725","7192775","726","1449","791","828","731","732","733","734","735","737","829","784","785","1453","795","796","798","1459","801","803","804","736","783","1452","790","1454","792","794","952","953","775")

# Gymnosperms

unique_depobio_species_htax %>%
mutate(group = case_when(
classKey == 359 ~ "Mammals",
classKey == 212 ~ "Birds",
classKey == 131 ~ "Amphibians",
classKey == 216 ~ "Insects",
orderKey %in% bony_fish ~ "Bony fish",
classKey %in% c("11493978","11418114","11569602","11592253") ~ "Reptiles",
phylumKey == 52 ~ "Molluscs",
classKey == 367 ~ "Arachnids",
phylumKey == 49 ~ "Flowering plants",
phylumKey == 35 ~ "Mosses",
phylumKey == 95 ~ "Sac fungi",
phylumKey == 34 ~ "Basidiomycota",
classKey == 7228684 ~ "Ferns",
TRUE ~ "other")) %>% 
group_by(group) %>%
summarise(n_occ = sum(count),
n_species = n_distinct(name)
) %>%
arrange(-n_occ) %>% 
knitr::kable()

```

## total unique literature citations 

```{r lit, echo=FALSE,include=FALSE}
lit_data = "https://api.gbif.org/v1/dataset/search/export?format=TSV&q=depobio&type=OCCURRENCE" %>%
readr::read_tsv() %>%
head(hh) %>%
select(dataset_key) %>%
glimpse() %>%  
mutate(lit_data = map(dataset_key,~rgbif::lit_search(datasetKey=.x,limit=1000)$data)) %>% 
tidyr::unnest(cols="lit_data")

# make exports 
readr::write_tsv(lit_data %>% select(-dataset_key) %>% unique(),"exports/lit_data.tsv")


```

The total number of literature citations for datasets using the keyword "depobio": 

`r lit_data %>% pull(id) %>% unique() %>% length()`

See full export of this data in `exports/lit_data.tsv`.


<!-- # number of records by department (to answer the question: are there records from départments and pays d’outre-mer outside “mainland” France?) -->
<!-- # gadm_gid=FRA.1_1&gadm_gid=FRA.2_1 -->

<!-- Not completed -->

<!-- ## impact assesments  -->
<!-- depbio eventDate n_occ, n_sp "about" and "published by" FR -->

<!-- eventDate means "first"  -->

<!-- make citations table for all of depobio as export  -->
